---
title: "data wrangling 2"
output: github_document
---

```{r}
library(tidyverse)
library(rvest)
library(httr)


knitr::opts_chunk$set(
  fig.width = 6, 
  fig.asp = .6, 
  out.width = "90%"
)
```

## Getting data from the web 

National Survey on Drug Use and Health (2013-2014 & 2014-2015)

```{r}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

drug_use_html = 
  read_html(url)

drug_use_df = 
  drug_use_html %>% 
  html_table() %>% 
  first() %>% 
  slice(-1)
```

The read_html function goes fully onto the internet every time. 

First row is the footnote (first row everywhere)

## Star Wars

Get some star wars data... 

```{r}
sw_url = "https://www.imdb.com/list/ls070150896/"

sw_html = 
  read_html(sw_url)
```

Not a well formatted thing, so we nee to use selector gadget 

```{r}
sw_titles = sw_html %>% 
  html_elements(".lister-item-header a") %>% 
  html_text()

sw_revenue = sw_html %>% 
  html_elements(".text-small:nth-child(7) span:nth-child(5)") %>% 
  html_text()

sw_df = 
  tibble(
    title = sw_titles, 
    revenue = sw_revenue
  )
```

## Napoleon Dynamite 

Dynamite reviews 

```{r}
dynamite_url = "https://www.amazon.com/product-reviews/B00005JNBQ/ref=cm_cr_arp_d_viewopt_rvwer?ie=UTF8&reviewerType=avp_only_reviews&sortBy=recent&pageNumber=1"

dynamite_html = 
  read_html(dynamite_url)

dynamite_review_titles = 
  dynamite_html %>% 
  html_elements(".a-text-bold span") %>% 
  html_text

dynamite_stars = 
  dynamite_html %>% 
  html_elements("#cm_cr-review_list .review-rating") %>% 
  html_text

dynamite_df = 
  tibble(titles = dynamite_review_titles, 
         stars = dynamite_stars)

```

## APIs 

Get some data from an API about water. 

```{r}
water_df = 
  GET("https://data.cityofnewyork.us/resource/ia2d-e54m.csv") %>% 
  content()
```

Let's see what JSON looks like...

```{r}
water_df = 
  GET("https://data.cityofnewyork.us/resource/ia2d-e54m.json") %>% 
  content("text") %>% 
  jsonlite::fromJSON() %>% 
  as_tibble()
```

## BRFSS data via API 

```{r}
brfss_df = 
  GET("https://chronicdata.cdc.gov/resource/acme-vg9e.csv",
      query = list("$limit" = 5000)) %>% 
  content() 
```

There are actuallly 134,000 rows, but it only gives the first 1,000. 
(Bandwidth protection)

## Pokemon API 

```{r}
pokemon_data = 
  GET("https://pokeapi.co/api/v2/pokemon/1") %>% 
  content()
  
pokemon_data[["name"]]
pokemon_data[["height"]]
pokemon_data[["abilities"]]
```

not nicely formatted as a csv

R package - wrappers for APIs 

rnoaa package 
rtweet (twitter data)








