---
title: "strings and factors"
output: github_document
---

```{r}
library(tidyverse)
library(rvest)
library(httr)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6, 
  fig.asp = .6, 
  out.width = "90%"
)


theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## String vectors

```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "m")

str_detect(string_vec, "jeff")

str_replace(string_vec, "jeff", "Jeff")

str_replace(string_vec, "jeff", "")

str_replace(string_vec, "e", "This is an uppercase E")
```

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "i think")

str_detect(string_vec, "i think$")
```

^ = the beggining of a line (first thing in the line)
$ = at the end, NOT the first thing in the line 

```{r}
string_vec = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
  )

str_detect(string_vec, "bush")

str_detect(string_vec, "Bush")

str_detect(string_vec, "[Bb]ush")
```

[] = upper or lowercase works

```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "[0-9][a-zA-Z]")
```

Everything 0-9 in the first digit and then any upper/ lowercase letters 

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
```

. = ANYTHING; 7 followed by literally anything and then 11 - that's what we want 
```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")
```

Square brackets and . have special meanings in regular expressions, so searching for those is a little tough 

If you're ACTUALLY looking for a square bracket, you need to add a backslash, but backslash is a special character, so you have to put 2 of them 

## Why factors are weird

```{r}
factor_vec = factor(c("male", "male", "female", "female"))
```

* you didn't tell me which was the reference cat, so it goes alphabetically 

```{r}
as.numeric(factor_vec)
```

* converts to underlying numbers 

```{r}
factor_vec = fct_relevel(factor_vec, "male")
```

## NSDUH

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj =
  read_html(nsduh_url) %>% 
  html_table() %>% 
  first() %>% 
  slice(-1)
```

* Presence of the a = character variable; lots of untidyness 

Let's clean it up

```{r}
table_marj %>% 
  select(-contains("P value")) %>% 
  pivot_longer(
    -State, 
    names_to = "age_year", 
    values_to = "percent") %>% 
  separate(age_year, into = c("age", "year"), -11)

marj_df = 
  table_marj %>% 
  select(-contains("P value")) %>% 
  pivot_longer(
    -State, 
    names_to = "age_year", 
    values_to = "percent") %>% 
  separate(age_year, into = c("age", "year"), "\\(") %>% 
  mutate(
    year = str_replace(year, "\\)", ""), 
    percent = str_replace(percent, "[a-c]$", ""), 
    percent = as.numeric(percent)
  ) %>% 
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```

Do dataframe stuff 

```{r}
marj_df %>% 
  filter(age == "12-17") %>% 
  mutate(
    State = fct_reorder(State, percent)
  ) %>% 
  ggplot(aes(x = State, y = percent, color = year)) + 
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

## Restaurant Inspections

```{r}
data("rest_inspec")
rest_inspec
```

rest_inspec %>% slice(1:100) %>%  view()
(how to look at things)

```{r}
rest_inspec %>% 
  janitor::tabyl(boro, grade)
```

```{r}
rest_inspec = 
  rest_inspec %>% 
  filter(
    str_detect(grade, "[ABC]"), 
   !(boro == "Missing")) %>% 
  mutate(boro = str_to_title(boro))

rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()

rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  mutate(
    boro = fct_infreq(boro)
  ) %>% 
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()
```

What about changing a label...

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  mutate(
    boro = fct_infreq(boro), 
    boro = str_replace(boro, "Manhattan", "The City")
  ) %>% 
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()
```

This worked, but undid the factor

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  mutate(
    boro = fct_infreq(boro), 
    boro = replace(boro, which(boro == "Manhattan"), "The City")
  ) %>% 
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()
```

If you're dealing with factors and you want to change the labels, it's annoying

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  mutate(
    boro = fct_infreq(boro), 
    boro = fct_recode(boro, "The City" = "Manhattan")
  ) %>% 
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()
```




